### 2011-06-10 thesis/novelty.tex
Finished explanation on how to detect novelty without assumption on P(novel) = constant.
No explanation on how to estimate the s_i parameters is given.

If on any graph structure the marginal probabilities of all variables except $a$
is the same either by using $A$ with $\phi$, or by using $A^\dag$ with $\phi^\dag$)
then $\sum_{A^\dag}\phi^\dag_i(a^\dag, \cdot) = s_i \sum_A \phi_i(a, \cdot)$.
It can be proved that since the distribution over any graph structure has to be
equal, then the same holds for a graph structure $P_{G_i}$ composed by a single
factor $\phi_i$:

\begin{eqnarray*}
P^\dag_{G_i}(x) &=& P_{G_i}(x) \\
\frac{\phi_i(x)}{Z^\dag_{G_i}}  &=& \frac{s_i \phi_i(x)}{Z_{G_i}} \\
s_i &=& \frac{Z_{G_i}}{Z^\dag_{G_i}}
\end{eqnarray*}

@@

\begin{eqnarray}
s &=& \frac{\Phi_G(x)}{\Phi^\dag_G(x)} \\
  &=& \frac{\sum_{a \in A}\prod \phi(a, x_\phi)}
           {\sum_{a \in A^\dag}\prod \phi^\dag(a^\dag, x_\phi)} \\
  &=& \frac{\sum_{a \in A}\prod \phi(a, x_\phi)}
           {\sum_{a \in A}\prod \phi(a, x_\phi) s_i} \\
  &=& \frac{1}{\prod s_i}
\end{eqnarray}

By adding a condition variable $c \in \{0, 1\}$, a condition factorthat is added result can be converted into a graph $G$is particularly interesting as it allows to detect novelty of $a$ by
having models to approach

Similar to the procedure followed before, a decision variable $c \in \{0, 1\}$
and a factor $\phi_c(c,a)$ will be added to $G$, creating graph $G'$.
If $\phi_c$ is defined as:

This is variable $c$ detects whether $a$ is a known or unknown class.
Note also that by adding this condition variable and factor any marginal
potential on $G$ is the same as the potential on $G'$. Since marginal potentials
can be seen as sum of potentials on all variables it can be proved by showing
the potential on any instance of all variables $X$ in $G$ is the same as the
marginal potential of the same variables in $G'$:

\begin{eqnarray*}
\Phi_{G'}(x) &=& \sum_{c}\prod_{\phi \in G'}{\phi(x_\phi)} \\
             &=& \left(\prod_{\phi \in G}{\phi(x_\phi)}\right)\phi_c(0,a) 
               + \left(\prod_{\phi \in G}{\phi(x_\phi)}\right)\phi_c(1,a)\\
             &=& \Phi_G(x)\phi_c(0,a) + \Phi_G(x)\phi_c(1,a)\\
             &=& \Phi_G(x)\\
\end{eqnarray*}

An optimal decision on whether variable $a$ of $G$ is $\in \{A_1,\dotsc,A_K\}$
can be performed by a map \gls{MAP} operation on it. And that decision can be
done solely based on potentials in $G'$:

\begin{eqnarray*}
P_G(a \in \{A_1,\dotsc,A_K\}|x) &<& P_G(a \in \{A_{K+1},\dotsc,A_N\}|x) \\
\Phi_{G'}(c = 0 \land x) &<& \Phi_{G'}(c = 1 \land x) \\
2 \Phi_{G'}(c = 0 \land x) &<& \Phi_{G'}(c=0 \land x) + \phi_{G'}(c = 1 \land x)\\
\end{eqnarray*}

By extending $a$ with duplicated states $\{A_1, \dotsc, A_K\}$, extending
the respective factors and adding a factor $\phi_f$ on the decision
variable $c$, it is possible to create a graph $G^{''}$ where a MAP decision on
$c$ will yield the same result. This graph $G^{''}$ is particullary 
the respective factors and , it is possible
to transform the previous condition on a map operation

And therefore the same happens to marginal potentials as marginal potentials
are the sum of potentials.

to detect whether a given sample from the graph has variable
$a$considered known to the system
The system only knowns $k$ of those $N$ states.
All the remaining ones are novel.

Just like in previous sections it was assumed that: it is possible to estimate
all the factors on the graph not dependent on $a$, using labelled data it is
possible to estimate the factors dependent on $a$ given that
$a \in \{A_1,\dotsc,A_K\}$. And using unlabelled data it is possible to estimate
the factors on the graph $P(x)$ for $a \in \{A_1,\dotsc,A_N\}$.

>>>>

If on any graph structure the marginal probabilities of all variables except $a$
is the same either by using $A$ with $\phi$, or by using $A^\dag$ with $\phi^\dag$
then $\sum_{A^\dag}\phi^\dag_i(a^\dag, \cdot) = s_i \sum_A \phi_i(a, \cdot)$.
It can be proved that since the distribution over any graph structure has to be
equal, then the same holds for a graph structure $P_{G_i}$ composed by a single
factor $\phi_i$:

\begin{eqnarray*}
P^\dag_{G_i}(x) &=& P_{G_i}(x) \\
\frac{\phi_i(x)}{Z^\dag_{G_i}}  &=& \frac{s_i \phi_i(x)}{Z_{G_i}} \\
s_i &=& \frac{Z_{G_i}}{Z^\dag_{G_i}} \quad \text{(constant)}
\end{eqnarray*}

@@

\begin{eqnarray*}
s &=& \frac{\Phi_G(x)}{\Phi^\dag_G(x)} \\
  &=& \frac{\sum_{a \in A}\prod \phi(a, x_\phi)}
           {\sum_{a \in A^\dag}\prod \phi^\dag(a^\dag, x_\phi)} \\
  &=& \frac{\sum_{a \in A}\prod \phi(a, x_\phi)}
           {\sum_{a \in A}\prod \phi(a, x_\phi) s_i} \\
  &=& \frac{1}{\prod s_i}
\end{eqnarray*}

If factor $\phi_i$ appears $k_i$ times on a graph structure then:
\begin{eqnarray}
\Phi^\dag_G(x) \prod {s_i}^{k_i} = \Phi_G(x)
\end{eqnarray}

Alternative, if factors $\phi^\ddag_i(a^\dag,x) = \phi^\dag_i(a^\dag,x)/s_i$ are
introduced then:
\begin{eqnarray}
\Phi^\ddag_G(x) = \Phi_G(x)
\end{eqnarray}

Having shown how to replace a variable $a$ for an equivalent $a^\dag$, and
respectively change the factors dependent on $a$ from $\phi_i(a,\cdot)$ to
$\phi^\ddag_i(a^\dag,\cdot)$ such that the potentials on the original
graph are kept constant it becomes possible to go back to the graph $G'$ and
replace the states $\{A^*_1,\dotsc,A^*_N\}$ for the equivalents $A^\ddag$
and change the factors associated with $a'$ for:

\begin{eqnarray*}
\phi_i(a', \cdot) &=& \left\{
  \begin{array}{l l}
    \phi_i(a',\cdot), & \qquad \text{if } a' \in \{A_1,\dotsc,A_K\} \\
    \phi^\ddag_i(a', \cdot), & \qquad \text{if } a' \in A^\ddag
  \end{array} \right. \\
\end{eqnarray*}

It is now possible to perform a decision on the novelty of $a$ with the condition:
\begin{equation}
2 \Phi_{G'}(x | c = 0) < \Phi_{G'}(x | c= 1)
\end{equation}

This condition is particularly useful since it does not depends on the normalization
factors of $G'$. This allows to use \emph{intractable models} for some factors,
e.g.\ using an \gls{SVM} for factors associated with sensed variables.

\subsection{Estimating the Factor Scaling-factors $s_i$}
On the previous section it was shown how to detect novelty on a variable $a$ by
approximating the unconditional distribution factors with factors $\phi^\dag_i$
that are a scalar $s_i$ away from the ones used on the real distribution.

This section address the problem of estimating those parameters.

\subsection{Old content}
